cardinalR: Generating interesting high-dimensional data structures

### 2025-05-04 paper-cardinalR.Rmd
Added cone description

@@

To generalize the Swiss roll structure to arbitrary dimensions, we introduce a function `generate_swiss_roll(n, p)`, which constructs a high-dimensional version of the classic 3D Swiss roll while preserving its core characteristics.

>>>>

To simulate a cone-shaped structure in arbitrary dimensions, we define a function `gen_cone(n, p, h, ratio)`, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.

This function generates `n` points in a `p`-dimensional space, where the last dimension (`X_p`) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.

Points along the height axis are drawn from an exponential distribution to increase density near the tip:

$$
X_p = h_i \sim \text{Exp}(\lambda = 2/h), \quad \text{truncated at } h
$$

The effective radius at height $h_i$ decreases linearly from base to tip, controlled by a shape ratio parameter:

$$
r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) \cdot \frac{h_i}{h}, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1
$$

For each point, a direction is sampled uniformly from a $(p-1)$-dimensional hypersphere using generalized spherical coordinates (angles). The radial coordinates are scaled by $r_i$, which ensures a conical taper. Specifically, when $p = 3$, this scaling results in a classic $3\text{-}D$ cone. However, for $p > 3$, the introduction of additional angular components allows for a smooth extension into higher dimensions, preserving the conical shape while accommodating the complexities of multi-dimensional geometry.

@@

To generalize the Swiss roll structure to arbitrary dimensions, we introduce a function `gen_swissroll(n, p, w)`, which constructs a high-dimensional version of the classic 3D Swiss roll while preserving its core characteristics.
